---
title: "homework3_rastermap"
author: "Fitrah Ramadhan"
date: "2023-10-18"
output: html_document
---
\\

### **PROBLEM STATEMENT**
1. The task is to manipulate some raster data and produce some descriptive statistics.\
2.1. For any country in the World: I choose United Kingdom
2.2. produce descriptive statistics that show the difference in maximum annual temperature for United Kingdom's key cities (delta tx) between SSP1 (SSP126) and SSP5 (SSP585) 
3.4. for the years 2081-2100
3.5. using any model and resolution (UKESM1).
\\

### **LOADING NECESSARY PACKAGES**
```{r}

library(tidyverse) #to do data frame/tibble wrangling
library(sf) #to do spatial data generation and manipulation
library(terra) #same as sf, but I think it is more specialized for raster data
library(raster) #same with terra
library(tibble) #convert dataframe to tibble
library(ggplot2)
library(plotly)
library(here)

```

### **UPLOAD DATA**
```{r}

#cities identity data
cities_map <- st_read(here("data/week3", "World_Cities.shp"))
cities_map <- cities_map %>%
  filter(str_detect(CNTRY_NAME, "United Kingdom")) %>%
  dplyr::select("CITY_NAME") 

#generate XY columns to extract tx in key cities
cities_coord <- as_tibble(st_coordinates(cities_map$geometry)) %>%
  dplyr::mutate(city = cities_map$CITY_NAME) %>%
  tibble::column_to_rownames(., var='city')

#united kingdom outline
st_layers(here("data/week3", "gadm41_GBR.gpkg"))
ukoutline_city <- st_read(here("data/week3", "gadm41_GBR.gpkg"), layer="ADM_ADM_4") %>% 
  group_by(NAME_2) %>% 
  summarize(geometry = st_union(geom))

#generate outline for uk
ukoutline_country <- ukoutline_city %>%
  summarize(st_union(geometry))

#generate outline for each cities
ukoutline_key_city <- ukoutline_city %>%
  filter(str_detect(NAME_2, paste(rownames(cities_coord), collapse="|")))

#world climate data
month_tmax <- c("tmax_Jan", 
           "tmax_Feb", 
           "tmax_Mar", 
           "tmax_Apr", 
           "tmax_May", 
           "tmax_Jun", 
           "tmax_Jul", 
           "tmax_Aug", 
           "tmax_Sep", 
           "tmax_Oct", 
           "tmax_Nov", 
           "tmax_Dec")

tmax_p126 <- terra::rast(here("data/week3", "wc2.1_2.5m_tmax_UKESM1-0-LL_ssp126_2081-2100.tif"))
names(tmax_p126) <- month_tmax

tmax_p585 <- terra::rast(here("data/week3", "wc2.1_2.5m_tmax_UKESM1-0-LL_ssp585_2081-2100.tif"))
names(tmax_p585) <- month_tmax
```

### **MANIPULATE CLIMATE DATA BY CITY (PIVOT WIDE): TO OBTAIN DELTA TX EACH MONTHS IN EACH CITIES**
```{r}

#UK Tx P126 in the wide form
UK_tmax_p126_month_wide <- terra::extract(tmax_p126,cities_coord) %>%
  dplyr::mutate(city = cities_map$CITY_NAME, .after=ID) %>%
  tibble::column_to_rownames(., var='city')

#UK Tx P585 in the wide form
UK_tmax_p585_month_wide <- terra::extract(tmax_p585,cities_coord) %>%
  dplyr::mutate(city = cities_map$CITY_NAME, .after=ID) %>%
  tibble::column_to_rownames(., var='city')

#Get the delta of Tx in each months in each cities
UK_tmax_delta_month <- round(as.data.frame(data.matrix(UK_tmax_p585_month_wide[2:13]) - data.matrix(UK_tmax_p126_month_wide[2:13])), digit=2)

```

### **MANIPULATE CLIMATE DATA BY CITY (PIVOT LONG): TO OBTAIN DELTA TX ANNUALLY IN EACH CITIES**
```{r}

#UK Tx P126 in the long form for each month
UK_tmax_p126_month_long <- terra::extract(tmax_p126,cities_coord) %>%
  dplyr::mutate(city = cities_map$CITY_NAME, .after=ID) %>%
  pivot_longer(cols=3:14, names_to= "Month", values_to = "tmax") 

#UK Tx P126 in the long form annually
UK_tmax_p126_annual <- UK_tmax_p126_month_long %>%
  group_by(city) %>%
  summarise(mean_tmax_p126 = round(mean(tmax), digit=3))

#UK Tx P585 in the long form for each month
UK_tmax_p585_month_long <- terra::extract(tmax_p585,cities_coord) %>%
  dplyr::mutate(city = cities_map$CITY_NAME, .after=ID) %>%
  pivot_longer(cols=3:14, names_to= "Month", values_to = "tmax") 

#UK Tx P585 in the long form annually
UK_tmax_p585_annual <- UK_tmax_p585_month_long %>%
  group_by(city) %>%
  summarise(mean_tmax_p585 = round(mean(tmax), digit=3))

#Get the delta of Tx annually in each cities
UK_tmax_delta_annual <- UK_tmax_p126_annual %>%
  left_join(., UK_tmax_p585_annual, by='city') %>%
  dplyr::mutate(delta_tmax = mean_tmax_p585 - mean_tmax_p126)

```

### **DESCRIPTIVE STATISTICS FOR KEY CITIES IN UK (COMPARISON)**
```{r}

#DEFINE FUNCTION TO GENERATE DESC_STATISTICS
desc_statistics_cities <- function(i) {
  
  ##PRODUCE MAPS
  city_tmax_p126_map <- mean(terra::crop(tmax_p126, ukoutline_key_city[i,]))
  city_tmax_p585_map <- mean(terra::crop(tmax_p585, ukoutline_key_city[i,]))
  city_tmax_diff_map <- city_tmax_p585_map - city_tmax_p126_map
  city_name <- as.character(ukoutline_key_city$NAME_2[i])
  title_string_diff <- paste("Map of Annual Delta Tx-P585 & Tx-P126 in", city_name, "for 2081-2100")
  plot(city_tmax_diff_map, main = title_string_diff)
  names(city_tmax_diff_map) <- city_name
  
  ##PRODUCE HISTOGRAM COMPARISON BETWEEN TX-P126 & TX-P585
  
  ###data
  tx_p126 <- as.data.frame(mean(terra::crop(tmax_p126, ukoutline_key_city[i,])))
  tx_p585 <- as.data.frame(mean(terra::crop(tmax_p585, ukoutline_key_city[i,])))
  data <- data.frame(tx_p126 = tx_p126$mean, tx_p585 = tx_p585$mean)
  
  ###Create overlay histograms with plotly
  histogram_cities <- plot_ly(data) %>%
    add_histogram(x = ~tx_p126, name = 'tx_p126', opacity = 0.5) %>%
    add_histogram(x = ~tx_p585, name = 'tx_p585', opacity = 0.5) %>%
    layout(
      title = paste("Histogram of Annual Delta Tx-P585 & Tx-P126 in", city_name, "for 2081-2100"),
      xaxis = list(title = "Temperature"),
      yaxis = list(title = "Frequency"),
      barmode = 'overlay')
  
  print(histogram_cities)
      
}

#PRINT THE DESCRIPTION FOR EACH CITIES
for (i in 1: nrow(ukoutline_key_city)) {desc_statistics_cities(i)}


```

### **DESCRIPTIVE STATISTICS IN UK (COMPARISON)**
```{r}

#PRODUCE DELTA MAP
UK_tmax_p126_map <- mean(terra::crop(tmax_p126, ukoutline_country))
UK_tmax_p585_map <- mean(terra::crop(tmax_p585, ukoutline_country))
UK_tmax_diff_map <- UK_tmax_p585_map - UK_tmax_p126_map
plot(UK_tmax_p126_map, main = paste("Map of Annual Tx-P126 in UK for 2081-2100"))
plot(UK_tmax_p585_map, main = paste("Map of Annual Tx-P585 in UK for 2081-2100"))
plot(UK_tmax_diff_map, main = paste("Map of Annual Delta Tx-P585 & Tx-P126 in UK for 2081-2100"))

#PRODUCE COMPARISON HISTOGRAM
##data
tx_p126 <- as.data.frame(mean(terra::crop(tmax_p126, ukoutline_country)))
tx_p585 <- as.data.frame(mean(terra::crop(tmax_p585, ukoutline_country)))
data_hist <- data.frame(tx_p126 = tx_p126$mean, tx_p585 = tx_p585$mean)
  
##Create overlay histograms with plotly
histogram_uk <- plot_ly(data_hist) %>%
  add_histogram(x = ~tx_p126, name = 'tx_p126', opacity = 0.5) %>%
  add_histogram(x = ~tx_p585, name = 'tx_p585', opacity = 0.5) %>%
  layout(
    title = paste("Histogram of Annual Delta Tx-P585 & Tx-P126 in UK for 2081-2100"),
    xaxis = list(title = "Temperature"),
    yaxis = list(title = "Frequency"),
    barmode = 'overlay')
  

print(histogram_uk)

```

## **DONE AND DUSTED! OUT OF THIS WORLD**
# INI BAHASA INDONESIA UNTUK KAMU: TERIMA KASIH SUDAH DATANG KEMARI
